---
title: "Data Science 1 Project"
author: "Julia Urh, Maria Christeson, Toria Davenport"
date: '`r format(Sys.time(), "%B %d, %Y @ %I:%M %p")`'
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Section 1: Jackpot! 

## What *is* the powerball lottery?:

* information or whatever


## Values of the Lottery Over Time: Background

+ Choices made

+ blah blah 

## Values of the Lottery Over Time

```{r, echo = FALSE}
#Maria's plots here. We have to include explanations for the choices we made as well. 
```


## Fairness and Independence of Winning Numbers: Background

+ blah

+ blah blah

## Fairness and Independence of Winning Numbers

```{r, echo = FALSE}
#Toria's plots here. We have to include explanations for the choices we made as well.
```


# Section 2: Population Ratios

## Background!

Any background info or blurbs Julia wants to say here

## Age ternary plots: Background

+ Choices made

+ blah blah 

## Age ternary plots

```{r, echo = FALSE}
#Julia's plots here. We have to include explanations for the choices we made as well.
```

## Dependency Ratios: Background

+ choices

+ blah

## Dependency Ratios: Background

```{r, echo = FALSE}
#Julia's other plots here 
```


# Conclusions