---
title: "Data Science 1 Project"
author: "Julia Urh, Maria Christeson, Toria Davenport"
date: '`r format(Sys.time(), "%B %d, %Y @ %I:%M %p")`'
output: 
  ioslides_presentation:
  theme: yeti
  highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

# Section 1: Jackpot! 

```{r, include=FALSE}
#Loading of datasets
library(dplyr)
library(XML)
library(stringr)
library(blscrapeR)
library(openxlsx)
library(tidyr)
library(lubridate)
library(readxl)
library(devtools)
library(ggtern)
library(gapminder)
library(gganimate)
library(ggplot2)
library(ggrepel)
library(tidyverse)
library(rvest)
library(magrittr)
library(ggmap)
library(stringr)

#loading winning numbers dataset
winningnumbers <- read.csv("https://data.ny.gov/api/views/d6yy-54nr/rows.csv?accessType=DOWNLOAD")

#winningnumbers

#loading jackpot vals
url <- "http://www.lottofactor.com/script/jackpot_history/draw_date/101"


```

```{r, include=FALSE}
jackpot_vals <- readHTMLTable(url)[[18]] %>% 
  mutate_all(funs(str_replace(., pattern = "Â", replacement = ""))) %>% 
  mutate_all(funs(str_trim(., "left")))

#renaming the variables
jackpot_vals <- rename(jackpot_vals, drawdate = `Â Draw Date`)

jackpot_vals <- rename(jackpot_vals, Change = `Â Change`)

jackpot_vals <- rename(jackpot_vals, Jackpot.Amount = `Â Jackpot Amount`)

# adjusting inflation
inflation_adj <- inflation_adjust(2019)

# editing winning numbers
winningnumbers <- winningnumbers %>% separate(Winning.Numbers, into = c("white1", "white2", "white3", "white4", "white5", "powerball"), sep = " ")

winningnumbers <- rename(winningnumbers, drawdate = Draw.Date)


#editing jackpot vals to begin to create new dataset
jackpot_vals <- jackpot_vals %>% separate("drawdate", into = c("drawday", "drawdate"), sep = " ")

jackpot_vals$Change = gsub("\\$", "", jackpot_vals$Change)
jackpot_vals$Change = gsub("Million", "", jackpot_vals$Change)

jackpot_vals$Jackpot.Amount = gsub("\\$", "", jackpot_vals$Jackpot.Amount)
jackpot_vals$Jackpot.Amount = gsub("Million", "", jackpot_vals$Jackpot.Amount)
jackpot_vals$Jackpot.Amount <- as.numeric(jackpot_vals$Jackpot.Amount)

jackpot_vals <- rename(jackpot_vals, jackpot_value = Jackpot.Amount)

#winningnumbers

#jackpot_vals

jackpot_vals$drawdate <- as.Date(jackpot_vals$drawdate, "%m/%d/%Y")
#jackpot_vals

winningnumbers$drawdate <- as.Date(winningnumbers$drawdate, "%m/%d/%Y")
#winningnumbers

jackpot_vals$drawdate <- ymd(jackpot_vals$drawdate)
jackpot_vals <- jackpot_vals %>%
  select(-Change)

#adding 2010 data
winningnumbers$drawdate <- ymd(winningnumbers$drawdate)
winningnumbers <- winningnumbers %>%
  select(-Multiplier)
drawdate <- c("2010-01-02", "2010-01-06", "2010-01-09", "2010-01-13", "2010-01-16", "2010-01-20", "2010-01-23", "2010-01-27", "2010-01-30")
white1 <- c(3, 20, 20, 8, 7, 15, 15, 6, 6)
white2 <- c(7, 27, 41, 11, 8, 32, 16, 15, 26)
white3 <- c(23, 29, 44, 15, 23, 43, 20, 20, 35)
white4 <- c(27, 56, 50, 32, 41, 44, 42, 26, 47)
white5 <- c(42, 58, 54, 50, 55, 46, 53, 31, 48)
powerball <- c(37, 31, 22, 27, 23, 19, 17, 9, 21)

powerball2 <- data.frame(drawdate, white1, white2, white3, white4, white5, powerball)
powerball2$drawdate <- as.Date(powerball2$drawdate)
powerball2$drawdate <- ymd(powerball2$drawdate)

winningnumbers$white1 <- as.numeric(winningnumbers$white1)
winningnumbers$white2 <- as.numeric(winningnumbers$white2)
winningnumbers$white3 <- as.numeric(winningnumbers$white3)
winningnumbers$white4 <- as.numeric(winningnumbers$white4)
winningnumbers$white5 <- as.numeric(winningnumbers$white5)
winningnumbers$powerball <- as.numeric(winningnumbers$powerball)
#winningnumbers

powerball1 <- rbind(winningnumbers, powerball2)
#powerball1

jackpot_vals <- jackpot_vals

powerball <- left_join(powerball1, jackpot_vals, by = "drawdate")
#powerball

powerball <- powerball %>%
  mutate(year = substr(drawdate, 1, 4))
#powerball


#adjusting inflation pt. 2
inflation_adj <- inflation_adjust(2019)

inflation_adj <- inflation_adj%>%
  select(-avg_cpi, -base_year, -pct_increase)

#inflation_adj

powerball <- left_join(powerball, inflation_adj, by = "year", match = "all")

powerballfinal <- powerball %>%
  mutate(jackpot_value_adj = (jackpot_value / adj_value))

#powerballfinal
```


## What *is* the powerball lottery?:

+ The numbers available  for the white balls were **below *50* from the start until approximately 2001**, **below *60* until 2014**, and became ***69* in 2015**. The powerball options show the opposite pattern, where they were **above *40* until 2009**, **above *30* until 2014**, and ***26* afterwards.**

+ In 2012, it appears that they added an increased starting jackpot at 40 Million dollars, and they added a 1 Million dollar prize for matching all 5 numbers. 


## Dataset:

```{r, echo = FALSE}
glimpse(powerballfinal)
```


## Values of the Lottery Over Time: Background

+ Objectives

    + Visualize patterns in money won over the years

    + Visualize winning values in 2019 dollars
    
+ Plots

    + Time series to show change
    
    + Interactive so that you can actually pinpoint specific datapoints 
    
    + Combined so that you can see difference in values (and inflation rate)

## Values of the Lottery Over Time

```{r, echo = FALSE}
# Library
library(dygraphs)
library(xts)          # To make the convertion data-frame / xts format
library(tidyverse)
library(lubridate)


# Then you can create the xts format, and thus use dygraph
don=xts(x = powerballfinal$jackpot_value, order.by = powerballfinal$drawdate)
don2=xts(x = powerballfinal$jackpot_value_adj, order.by = powerballfinal$drawdate)

don3 <- cbind(don, don2)

dygraph(don3) %>%
  dySeries("don", label = "Non-Adjusted") %>%
  dySeries("don2", label = "Adjusted") %>%
  dyOptions(stackedGraph = TRUE) %>%
  dyRangeSelector(height = 20)
```


## Fairness and Independence of Winning Numbers: Background

+ Objectives

    + Detect patterns in winning numbers
    
    + Compare the winning numbers over time 
    
+ Plots

    + Scatterplots to show basic patterns with time color-coded
    
    + Density plots for the patterns most easily detected in the scatterplot
    
    + Combined density plot to show differences between/relationships with the winning numbers for different balls

## Fairness and Independence of Winning Numbers

```{r, echo = FALSE}
library(ggplot2)

ggplot(powerballfinal, aes(x = drawdate, y = white1, col = drawdate)) +
  geom_point() +
  theme_classic()
```

## Fairness and Independence of Winning Numbers

```{r, echo = FALSE}
ggplot(powerballfinal, aes(x = drawdate, y = white5, col = drawdate)) +
  geom_point() +
  theme_classic()
```

## Fairness and Independence of Winning Numbers

```{r, echo = FALSE}
p1 <- ggplot(powerballfinal, aes(x = white5, col = drawdate)) +
  geom_density(aes(col = drawdate)) +
  theme_classic()

p1+ geom_vline(aes(xintercept=mean(white5)),
            color="blue", linetype="dashed", size=1) +
    geom_text(aes(x = mean(white5), y = 0.031, label = "White 5 mean"), color = "black", size = 4) +
  geom_vline(aes(xintercept = 60), color = "purple", linetype = "dashed", size =1) +
  labs(caption = "*Purple line indicates where maximum value was up until 2015") +
  xlab("White 5")
```


## Fairness and Independence of Winning Numbers

```{r, echo = FALSE}
p2 <- ggplot(powerballfinal, aes(x = white1, fill = drawdate)) +
  geom_density() +
  theme_classic()

p2+ geom_vline(aes(xintercept=mean(white1)),
            color="blue", linetype="dashed", size=1) +
  xlab("White 1")

```


## Fairness and Independence of Winning Numbers

```{r, echo = FALSE}
library(ggridges)

likelihood <- ggplot(powerballfinal) + 
  geom_density(aes(x = white1), color = "aquamarine4") +
  geom_density(aes(x = white2), color = "purple") +
  geom_density(aes(x = white3), color = "darkred") + 
  geom_density(aes(x = white4), color = "green") +
  geom_density(aes(x = white5)) +
  geom_density(aes(x = powerball), color = "deeppink2") +
  labs(x = "Ball value", y = "Density", title = "Distribution of Winning Numbers", subtitle = "Based off of balltype; Year > 2009") +
  geom_vline(aes(xintercept=mean(white1)),
            color="aquamarine4", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white1), y = 0.041, label = "White 1"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white2)),
            color="purple", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white2), y = 0.041, label = "White 2"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white3)),
            color="darkred", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white3), y = 0.041, label = "White 3"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white4)),
            color="green", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white4), y = 0.041, label = "White 4"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white5)),
            color="black", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white5), y = 0.041, label = "White 5"), color = "black" ) +
  geom_vline(aes(xintercept=mean(powerball)),
            color="deeppink2", linetype="dashed", size= 0.25) +
  geom_text(aes(x = mean(powerball), y = 0.021, label = "Powerball"), color = "black" ) +
  theme_classic() 

likelihood
```

## Fairness and Independence of Winning Numbers


Upon further investigiation, it appears that the white balls are always reported from highest to lowest. In light of this, it looks like there is indeed an equally likely chance of choosing any ball, keeping in mind the recently increased maximum. Combined, these bars present a fairly even spread of the values 0-69


## Fairness and Independence of Winning Numbers 


```{r, echo = FALSE}
library(ggridges)

likelihood <- ggplot(powerballfinal) + 
  labs(x = "Ball value", y = "Density", title = "Distribution of Winning Numbers", subtitle = "Based off of balltype; Year > 2009") +
  geom_vline(aes(xintercept=mean(white1)),
            color="aquamarine4", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white1), y = 0.041, label = "White 1"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white2)),
            color="purple", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white2), y = 0.041, label = "White 2"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white3)),
            color="darkred", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white3), y = 0.041, label = "White 3"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white4)),
            color="green", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white4), y = 0.041, label = "White 4"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white5)),
            color="black", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white5), y = 0.041, label = "White 5"), color = "black" ) +
  geom_vline(aes(xintercept=mean(powerball)),
            color="deeppink2", linetype="dashed", size= 0.25) +
  geom_text(aes(x = mean(powerball), y = 0.021, label = "Powerball"), color = "black" ) +
  theme_classic() 

likelihood 
```

# Section 2: Population Ratios

```{r, include=FALSE}
url <- "https://stat-jet-asu.github.io/DataScience1Private/datasets/gapminder_pop_data.xlsx"
destfile <- "datasets.xlsx"
download.file(url, destfile)
pop_0_4 <- read_excel(destfile, sheet = 1)
pop_5_9 <- read_excel(destfile, sheet = 2)
pop_10_14 <- read_excel(destfile, sheet = 3)
pop_15_19 <- read_excel(destfile, sheet = 4)
pop_20_39 <- read_excel(destfile, sheet = 5)
pop_40_59 <- read_excel(destfile, sheet = 6)
pop_60_100 <- read_excel(destfile, sheet = 7)
pop_tot_pop <- read_excel(destfile, sheet = 8)
```

## Background!

For this section we took the population ratios for most of the countries across the world and made them into a consumable format.

## Age ternary plots: Background

+ To create these plots, all the data had to come together into one dataframe. 

## A look at the population dataframe

```{r, include = FALSE}
population_0_4 <- rename(pop_0_4, "country" = "Total 0-4 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_0_4" = "population")
population_5_9 <- rename(pop_5_9, "country" = "Total 5-9 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_5_9" = "population")
population_10_14 <- rename(pop_10_14, "country" = "Total 10-14 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_10_14" = "population")

population_0_14 <- semi_join(population_0_4, population_5_9, population_10_14, by = c("country", "year")) %>%
  mutate(population_0_4$pop_0_4 + population_5_9$pop_5_9 + population_10_14$pop_10_14) %>%
  rename("pct0to14" = "population_0_4$pop_0_4 + population_5_9$pop_5_9 + population_10_14$pop_10_14") %>%
  select(-pop_0_4)

population_15_19 <- rename(pop_15_19, "country" = "Total 15-19 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_15_19" = "population")
population_20_39 <- rename(pop_20_39, "country" = "Total 20-39 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_20_39" = "population")
population_40_59 <- rename(pop_40_59, "country" = "Total 40-59 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_40_59" = "population")

population_15_59 <- semi_join(population_15_19, population_20_39, population_40_59, by = c("country", "year")) %>%
  mutate(population_15_19$pop_15_19 + population_20_39$pop_20_39 + population_40_59$pop_40_59) %>%
  rename("pct15to59" = "population_15_19$pop_15_19 + population_20_39$pop_20_39 + population_40_59$pop_40_59") %>%
  select(-pop_15_19)

population_60_plus <- rename(pop_60_100, "country" = "Total above 60 (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pct60plus" = "population")

population_0_59 <- full_join(population_0_14, population_15_59, by = c("country", "year"))

population <- full_join(population_0_59, population_60_plus, by = c("country", "year")) %>%
  mutate(pct0to14 / pct15to59) %>%
  rename("cdr" = "pct0to14/pct15to59") %>%
  mutate(pct60plus / pct15to59) %>%
  rename("odr" = "pct60plus/pct15to59") %>%
  mutate((pct0to14 + pct60plus) / pct15to59) %>%
  rename("tdr" = "(pct0to14 + pct60plus)/pct15to59")

```

```{r}
glimpse(population)
```

## Age ternary plots

```{r, echo = FALSE}
population_usa <- filter(population, country == "United States")
ggtern(population_usa, aes(x = pct0to14, y = pct15to59, z = pct60plus)) +
  geom_point(aes(color = year, shape = year < 2020), size = 3) +
  scale_shape_manual(values=c(1, 20),
                     labels = c("Predicted", "Recorded")) +
  ggtitle("Age Groups Over Time In USA") +
  labs(x = "0 to 14",
       y = "15 to 59",
       z = "60 plus") +
  theme_zoom_R() +
  theme(legend.justification = c(1, 1)) +
  labs(shape = "Prediction vs. Recorded", colour = "Year") 
```  

## Age ternary plots

```{r, echo = FALSE}
population_usa <- filter(population, country == "Slovenia")
ggtern(population_usa, aes(x = pct0to14, y = pct15to59, z = pct60plus)) +
  geom_point(aes(color = year, shape = year < 2020), size = 3) +
  scale_shape_manual(values=c(1, 20),
                     labels = c("Predicted", "Recorded")) +
  ggtitle("Age Groups Over Time In Slovania") +
  labs(x = "0 to 14",
       y = "15 to 59",
       z = "60 plus") +
  theme_zoom_R() +
  theme(legend.justification = c(1, 1)) +
  labs(shape = "Prediction vs. Recorded", colour = "Year")
```

## Dependency Ratios: Background

+ For the dependency ratios, we just wanted to look at the predicted year 2020.

+ There were a couple places on the original population dataframe that were not really countries, so they were dropped. (Channel Islands; 	Hong Kong, China; Macao, China; Netherlands Antilles; West Bank and Gaza)

+ You will also find a very obvious missing country on the maps and that is because South Sudan was not founded until 2011

## Dependency Ratios: 

```{r, include=FALSE}
map.world <- map_data("world")
population_2020 <- population %>%
  filter(year == 2020)                       
map.world.c <- map.world %>%
  rename("country" = "region")
joined_country <- semi_join(population_2020, map.world.c, by = "country")
j_country <- pull(joined_country, country)


pop_2020_good <- population_2020 %>%
  mutate(country = case_when(country == 'Central African Rep.' ~ 'Central African Republic',
                             country == 'Congo, Dem. Rep.' ~ 'Democratic Republic of the Congo',
                             country == 'Congo, Rep.' ~ 'Republic of Congo',
                             country == "Cote d'Ivoire" ~ "Ivory Coast",
                             country == "Czech Rep." ~ "Czech Republic",
                             country == "Dominican Rep." ~ "Dominican Republic",
                             country == "Korea, Dem. Rep." ~ "North Korea",
                             country == "Korea, Rep." ~ "South Korea",
                             country == "Macedonia, FYR" ~ "Macedonia",
                             country == "Micronesia, Fed. Sts." ~ "Micronesia",
                             country == "Saint Vincent and the Grenadines" ~ "Saint Vincent",
                             country == "Slovak Republic" ~ "Slovakia",
                             country == "Trinidad and Tobago" ~ "Trinidad",
                             country == "United Kingdom" ~ "UK",
                             country == "United States" ~ "USA",
                             country == "Virgin Islands (U.S.)" ~ "Virgin Islands",
                             country == "Yemen, Rep." ~ "Yemen",
                             country %in% (j_country) ~ country
                             )) %>%
  filter(country != is.na(country))
```


```{r, echo = FALSE}
map.world_joined <- left_join(map.world.c, pop_2020_good, by = 'country')
ggplot() +
  geom_polygon(data = map.world_joined, aes(x = long, y = lat, group = group, fill = tdr)) +
  theme_void() + 
  labs(title = 'Total Dependancy Ratios of World Countries') +
  theme(text = element_text(family = "Gill Sans", color = "#FFFFFF")
        ,panel.background = element_rect(fill = "#444444")
        ,plot.background = element_rect(fill = "#444444")
        ,panel.grid = element_blank()
        ,plot.title = element_text(size = 20)
        ,plot.subtitle = element_text(size = 10)
        ,axis.text = element_blank()
        ,axis.title = element_blank()
        ,axis.ticks = element_blank()
        ) 
```

## Dependency Ratios: 

```{r, echo = FALSE}
ggplot() +
  geom_polygon(data = map.world_joined, aes(x = long, y = lat, group = group, fill = cdr)) +
  theme_void() + 
  labs(title = 'Child Dependancy Ratios of World Countries') +
  theme(text = element_text(family = "Gill Sans", color = "#FFFFFF")
        ,panel.background = element_rect(fill = "#444444")
        ,plot.background = element_rect(fill = "#444444")
        ,panel.grid = element_blank()
        ,plot.title = element_text(size = 20)
        ,plot.subtitle = element_text(size = 10)
        ,axis.text = element_blank()
        ,axis.title = element_blank()
        ,axis.ticks = element_blank()
        ) +
  annotate("text", x = 20, y = -30, label = "watch Africa change", color = "white") +
  annotate("text", x = -100, y = 50, label = "watch America change", color = "white")
```

## Dependency Ratios: 

```{r, echo = FALSE}
ggplot() +
  geom_polygon(data = map.world_joined, aes(x = long, y = lat, group = group, fill = odr)) +
  theme_void() + 
  labs(title = 'Older Dependancy Ratios of World Countries') +
  theme(text = element_text(family = "Gill Sans", color = "#FFFFFF")
        ,panel.background = element_rect(fill = "#444444")
        ,plot.background = element_rect(fill = "#444444")
        ,panel.grid = element_blank()
        ,plot.title = element_text(size = 20)
        ,plot.subtitle = element_text(size = 10)
        ,axis.text = element_blank()
        ,axis.title = element_blank()
        ,axis.ticks = element_blank()
        
        )
```

# Questions?

```{r, include = FALSE}
sessionInfo()
```

