---
title: 'Part 1 Visual 2'
author: ''
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_float: false
    toc_depth: 4
    code_folding: hide
---

```{r setup_chunk, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = "")
```


# Code for powerball dataset

Packages:

```{r}
library(dplyr)
library(XML)
library(stringr)
library(blscrapeR)
library(openxlsx)
library(tidyr)
library(lubridate)

```


Initial Loading of Powerball dataset:

```{r}
#loading winning numbers dataset
winningnumbers <- read.csv("https://data.ny.gov/api/views/d6yy-54nr/rows.csv?accessType=DOWNLOAD")

#winningnumbers

#loading jackpot vals
url <- "http://www.lottofactor.com/script/jackpot_history/draw_date/101"

jackpot_vals <- readHTMLTable(url)[[18]] %>% 
  mutate_all(funs(str_replace(., pattern = "Â", replacement = ""))) %>% 
  mutate_all(funs(str_trim(., "left")))

#renaming the variables
jackpot_vals <- rename(jackpot_vals, drawdate = `Â Draw Date`)

jackpot_vals <- rename(jackpot_vals, Change = `Â Change`)

jackpot_vals <- rename(jackpot_vals, Jackpot.Amount = `Â Jackpot Amount`)

# adjusting inflation
inflation_adj <- inflation_adjust(2019)

# editing winning numbers
winningnumbers <- winningnumbers %>% separate(Winning.Numbers, into = c("white1", "white2", "white3", "white4", "white5", "powerball"), sep = " ")

winningnumbers <- rename(winningnumbers, drawdate = Draw.Date)


#editing jackpot vals to begin to create new dataset
jackpot_vals <- jackpot_vals %>% separate("drawdate", into = c("drawday", "drawdate"), sep = " ")

jackpot_vals$Change = gsub("\\$", "", jackpot_vals$Change)
jackpot_vals$Change = gsub("Million", "", jackpot_vals$Change)

jackpot_vals$Jackpot.Amount = gsub("\\$", "", jackpot_vals$Jackpot.Amount)
jackpot_vals$Jackpot.Amount = gsub("Million", "", jackpot_vals$Jackpot.Amount)
jackpot_vals$Jackpot.Amount <- as.numeric(jackpot_vals$Jackpot.Amount)

jackpot_vals <- rename(jackpot_vals, jackpot_value = Jackpot.Amount)

#winningnumbers

#jackpot_vals

```

Creating dataset "powerball"

```{r}
jackpot_vals$drawdate <- as.Date(jackpot_vals$drawdate, "%m/%d/%Y")
#jackpot_vals

winningnumbers$drawdate <- as.Date(winningnumbers$drawdate, "%m/%d/%Y")
#winningnumbers

jackpot_vals$drawdate <- ymd(jackpot_vals$drawdate)
jackpot_vals <- jackpot_vals %>%
  select(-Change)


winningnumbers$drawdate <- ymd(winningnumbers$drawdate)
winningnumbers <- winningnumbers %>%
  select(-Multiplier)




```


```{r}
drawdate <- c("2010-01-02", "2010-01-06", "2010-01-09", "2010-01-13", "2010-01-16", "2010-01-20", "2010-01-23", "2010-01-27", "2010-01-30")
white1 <- c(3, 20, 20, 8, 7, 15, 15, 6, 6)
white2 <- c(7, 27, 41, 11, 8, 32, 16, 15, 26)
white3 <- c(23, 29, 44, 15, 23, 43, 20, 20, 35)
white4 <- c(27, 56, 50, 32, 41, 44, 42, 26, 47)
white5 <- c(42, 58, 54, 50, 55, 46, 53, 31, 48)
powerball <- c(37, 31, 22, 27, 23, 19, 17, 9, 21)

powerball2 <- data.frame(drawdate, white1, white2, white3, white4, white5, powerball)
powerball2$drawdate <- as.Date(powerball2$drawdate)
powerball2$drawdate <- ymd(powerball2$drawdate)

winningnumbers$white1 <- as.numeric(winningnumbers$white1)
winningnumbers$white2 <- as.numeric(winningnumbers$white2)
winningnumbers$white3 <- as.numeric(winningnumbers$white3)
winningnumbers$white4 <- as.numeric(winningnumbers$white4)
winningnumbers$white5 <- as.numeric(winningnumbers$white5)
winningnumbers$powerball <- as.numeric(winningnumbers$powerball)
#winningnumbers

powerball1 <- rbind(winningnumbers, powerball2)
#powerball1
```


#adding 2010 data

```{r}
jackpot_vals <- jackpot_vals

powerball <- left_join(powerball1, jackpot_vals, by = "drawdate")
#powerball
```


```{r}

powerball <- powerball %>%
  mutate(year = substr(drawdate, 1, 4))
#powerball
```

```{r}
inflation_adj <- inflation_adjust(2019)

inflation_adj <- inflation_adj%>%
  select(-avg_cpi, -base_year, -pct_increase)

#inflation_adj
```

```{r}
powerball <- left_join(powerball, inflation_adj, by = "year", match = "all")

powerballfinal <- powerball %>%
  mutate(jackpot_value_adj = (jackpot_value / adj_value))

#powerballfinal
```



**The numbers available were below 50 from the start until approximately 2001, below 60 until 2014, and became 69 in 2015. The powerball options show the opposite pattern, where they were above 40 until 2009, above 30 until 2014, and 26 afterwards.**

### White 1, numbers range currently from 1-69.

```{r}
library(ggplot2)

ggplot(powerballfinal, aes(x = drawdate, y = white1, col = drawdate)) +
  geom_point()
```

### White 2, numbers range currently from 1-69

```{r}
library(ggplot2)

ggplot(powerballfinal, aes(x = drawdate, y = white2, col = drawdate)) +
  geom_point()
```


### White 3, numbers range currently from 1-69

```{r}
library(ggplot2)

ggplot(powerballfinal, aes(x = drawdate, y = white3, col = drawdate)) +
  geom_point()
```


### White 4, numbers range currently from 1-69

```{r}
library(ggplot2)

ggplot(powerballfinal, aes(x = drawdate, y = white4, col = drawdate)) +
  geom_point()
```


### White 5, numbers range currently from 1-69


```{r}
library(ggplot2)

ggplot(powerballfinal, aes(x = drawdate, y = white5, col = drawdate)) +
  geom_point()
```

### Powerball, numbers currently ranging from 1-29.

```{r}
library(ggplot2)

ggplot(powerballfinal, aes(x = drawdate, y = powerball, col = drawdate)) +
  geom_point()
```




**After intitial analysis, white 5 and white 1 look the most unevenly distributed. White 1 is more concentrated at the low end, while white 5 is more concentrated at the high end.**


### White 5

```{r}
library(ggplot2)

p1 <- ggplot(powerballfinal, aes(x = white5, col = drawdate)) +
  geom_density(aes(col = drawdate))

p1+ geom_vline(aes(xintercept=mean(white5)),
            color="blue", linetype="dashed", size=1)
```



### White 1

```{r}
library(ggplot2)

p2 <- ggplot(powerballfinal, aes(x = white1, fill = drawdate)) +
  geom_density()

p2+ geom_vline(aes(xintercept=mean(white1)),
            color="blue", linetype="dashed", size=1)
```







# Plot of All Distributions

```{r}
library(ggridges)

likelihood <- ggplot(powerballfinal) + 
  geom_density(aes(x = white1), color = "aquamarine4") +
  geom_density(aes(x = white2), color = "purple") +
  geom_density(aes(x = white3), color = "darkred") + 
  geom_density(aes(x = white4), color = "green") +
  geom_density(aes(x = white5)) +
  geom_density(aes(x = powerball), color = "deeppink2") +
  labs(x = "Ball value", y = "Density", title = "Distribution of Winning Numbers", subtitle = "Based off of balltype; Year > 2009") +
  geom_vline(aes(xintercept=mean(white1)),
            color="aquamarine4", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white1), y = 0.041, label = "White 1"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white2)),
            color="purple", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white2), y = 0.041, label = "White 2"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white3)),
            color="darkred", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white3), y = 0.041, label = "White 3"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white4)),
            color="green", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white4), y = 0.041, label = "White 4"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white5)),
            color="black", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white5), y = 0.041, label = "White 5"), color = "black" ) +
  geom_vline(aes(xintercept=mean(powerball)),
            color="deeppink2", linetype="dashed", size= 0.25) +
  geom_text(aes(x = mean(powerball), y = 0.021, label = "Powerball"), color = "black" ) +
  theme_classic() 

likelihood

```


```{r}
likelihood <- ggplot(powerballfinal) +
  labs(x = "Ball value", y = "Density", title = "Distribution of Winning Numbers", subtitle = "Based off of balltype; Year > 2009") +
  geom_vline(aes(xintercept=mean(white1)),
            color="aquamarine4", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white1), y = 0.041, label = "White 1"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white2)),
            color="purple", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white2), y = 0.041, label = "White 2"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white3)),
            color="darkred", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white3), y = 0.041, label = "White 3"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white4)),
            color="green", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white4), y = 0.041, label = "White 4"), color = "black" ) +
  geom_vline(aes(xintercept=mean(white5)),
            color="black", linetype="dashed", size=0.5) +
  geom_text(aes(x = mean(white5), y = 0.041, label = "White 5"), color = "black" ) +
  geom_vline(aes(xintercept=mean(powerball)),
            color="deeppink2", linetype="dashed", size= 0.25) +
  geom_text(aes(x = mean(powerball), y = 0.021, label = "Powerball"), color = "black" ) +
  theme_classic() 

likelihood
```



