---
title: 'Project 2 Pop Ratio'
author: ''
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_float: false
    toc_depth: 4
    code_folding: hide
---

```{r setup_chunk, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = "")
```

Packages:

```{r}
library(dplyr)
library(XML)
library(stringr)
library(ggplot2)
library(readxl)
```

#Population Ratios
```{r}
url <- "https://stat-jet-asu.github.io/DataScience1Private/datasets/gapminder_pop_data.xlsx"
destfile <- "datasets.xlsx"
download.file(url, destfile)
pop_0_4 <- read_excel(destfile, sheet = 1)
pop_5_9 <- read_excel(destfile, sheet = 2)
pop_10_14 <- read_excel(destfile, sheet = 3)
pop_15_19 <- read_excel(destfile, sheet = 4)
pop_20_39 <- read_excel(destfile, sheet = 5)
pop_40_59 <- read_excel(destfile, sheet = 6)
pop_60_100 <- read_excel(destfile, sheet = 7)
pop_tot_pop <- read_excel(destfile, sheet = 8)
```
Create a single dataset called population, with the following variable structure. Verify this using glimpse() on the dataset in your report.
country (chr) – country name
year (num) – year
pct0to14 (num) – percentage of population age 0 to 14
pct15to59 (num) – percentage of population age 15 to 59
pct60plus (num) – percentage of population age 60+
cdr (num) – child dependency ratio (age 0-14 / age 15-59)
odr (num) – old-age dependency ratio (age 60+ / age 15-59)
tdr (num) – total dependency ratio (age 0-19 + age 60+ / age 15-59)

```{r}
population <- rename(pop_tot_pop, "country" = "Total population")
  
```