---
title: 'Project 2 Pop Ratio'
author: ''
date: '`r format(Sys.time(), "%A, %B %d, %Y @ %X")`'
output: 
  html_document: 
    theme: yeti
    highlight: textmate
    toc: true
    toc_float: false
    toc_depth: 4
    code_folding: hide
---

```{r setup_chunk, echo = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE, comment = "")
devtools::install_github('thomasp85/gganimate', force = TRUE)
```

Packages:

```{r}
library(dplyr)
library(XML)
library(stringr)
library(ggplot2)
library(readxl)
library(tidyr)
library(devtools)
library(ggtern)
library(gapminder)
library(gganimate)
library(ggplot2)
library(ggrepel)
```

#Population Ratios
```{r}
url <- "https://stat-jet-asu.github.io/DataScience1Private/datasets/gapminder_pop_data.xlsx"
destfile <- "datasets.xlsx"
download.file(url, destfile)
pop_0_4 <- read_excel(destfile, sheet = 1)
pop_5_9 <- read_excel(destfile, sheet = 2)
pop_10_14 <- read_excel(destfile, sheet = 3)
pop_15_19 <- read_excel(destfile, sheet = 4)
pop_20_39 <- read_excel(destfile, sheet = 5)
pop_40_59 <- read_excel(destfile, sheet = 6)
pop_60_100 <- read_excel(destfile, sheet = 7)
pop_tot_pop <- read_excel(destfile, sheet = 8)
```

```{r}
population_0_4 <- rename(pop_0_4, "country" = "Total 0-4 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_0_4" = "population")
population_5_9 <- rename(pop_5_9, "country" = "Total 5-9 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_5_9" = "population")
population_10_14 <- rename(pop_10_14, "country" = "Total 10-14 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_10_14" = "population")

population_0_14 <- semi_join(population_0_4, population_5_9, population_10_14, by = c("country", "year")) %>%
  mutate(population_0_4$pop_0_4 + population_5_9$pop_5_9 + population_10_14$pop_10_14) %>%
  rename("pct0to14" = "population_0_4$pop_0_4 + population_5_9$pop_5_9 + population_10_14$pop_10_14") %>%
  select(-pop_0_4)

population_15_19 <- rename(pop_15_19, "country" = "Total 15-19 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_15_19" = "population")
population_20_39 <- rename(pop_20_39, "country" = "Total 20-39 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_20_39" = "population")
population_40_59 <- rename(pop_40_59, "country" = "Total 40-59 years (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pop_40_59" = "population")

population_15_59 <- semi_join(population_15_19, population_20_39, population_40_59, by = c("country", "year")) %>%
  mutate(population_15_19$pop_15_19 + population_20_39$pop_20_39 + population_40_59$pop_40_59) %>%
  rename("pct15to59" = "population_15_19$pop_15_19 + population_20_39$pop_20_39 + population_40_59$pop_40_59") %>%
  select(-pop_15_19)

population_60_plus <- rename(pop_60_100, "country" = "Total above 60 (%)") %>%
  gather("year", "population", 2:22) %>%
  rename("pct60plus" = "population")

population_0_59 <- full_join(population_0_14, population_15_59, by = c("country", "year"))

population <- full_join(population_0_59, population_60_plus, by = c("country", "year")) %>%
  mutate(pct0to14 / pct15to59) %>%
  rename("cdr" = "pct0to14/pct15to59") %>%
  mutate(pct60plus / pct15to59) %>%
  rename("odr" = "pct60plus/pct15to59") %>%
  mutate((pct0to14 + pct60plus) / pct15to59) %>%
  rename("tdr" = "(pct0to14 + pct60plus)/pct15to59")

glimpse(population)
```

```{r}
population_usa <- filter(population, country == "United States")
ggtern(population_usa, aes(x = pct0to14, y = pct15to59, z = pct60plus)) +
  geom_point(aes(color = year, shape = year < 2020), size = 3) +
  scale_shape_manual(values=c(1, 20),
                     labels = c("Predicted", "Recorded")) +
  ggtitle("Age Groups Over Time In USA") +
  labs(x = "0 to 14",
       y = "15 to 59",
       z = "60 plus") +
  theme_zoom_R() +
  theme(legend.justification = c(1, 1)) +
  labs(shape = "Prediction vs. Recorded", colour = "Year") 
  
  
population_usa <- filter(population, country == "Slovenia")
ggtern(population_usa, aes(x = pct0to14, y = pct15to59, z = pct60plus)) +
  geom_point(aes(color = year, shape = year < 2020), size = 3) +
  scale_shape_manual(values=c(1, 20),
                     labels = c("Predicted", "Recorded")) +
  ggtitle("Age Groups Over Time In Slovania") +
  labs(x = "0 to 14",
       y = "15 to 59",
       z = "60 plus") +
  theme_zoom_R() +
  theme(legend.justification = c(1, 1)) +
  labs(shape = "Prediction vs. Recorded", colour = "Year")
  
```

```{r}
letters <- c("A", "A", "A", "A", "A", "A", "A", "A", "A", "A",
             "B", "B", "B", "B", "B", "B", "B", "B", "B", "B", "B", "B", "B", "B", "B", "B", "B", 
             "C", "C", "C", "C", "C", "C", "C", "C", "C", "C", "C", "C", "C", "C", "C", "C", "C", "C", "C", 
             "D", "D", "D",
             "E", "E", "E", "E", "E", "E", "E",
             "F", "F", "F", "F", "F", 
             "G", "G", "G", "G", "G", "G", "G", "G", "G", "G", "G", "G", "G", 
             "H", "H", "H", "H",
             "I", "I", "I", "I", "I", "I", "I", "I", 
             "J", "J", "J", 
             "K", 'K', 'K', 'K', 'K', 'K', 
             'L', 'L', 'L', 'L', 'L', 'L', 'L', 'L', 
             'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M', 'M',
             'N', 'N', 'N', 'N', 'N', 'N', 'N', 'N', 'N', 'N',
             'O',
             'P', 'P', 'P', 'P', 'P', 'P', 'P', 'P', 'P', 
             'Q',
             'R', 'R', 'R', 'R',
             'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 'S', 
             'T', 'T', 'T', 'T', 'T', 'T', 'T', 'T', 'T', 'T', 
             'U', 'U', 'U', 'U', 'U', 'U', 'U', 
             'V', 'V', 'V', 'V', 
             'W', 'W',
             'Y',
             'Z', 'Z')
letter <- gsub('', "", letters) %>%
  as.character()
```

Create one or more visualizations to display and compare the projected 2020 dependency ratios for all countries. Use annotations or some other appropriate method to highlight at least one feature in the data you think might be interesting or noteworthy to an audience concerned about these issues. You may research other variables of interest to add to your visualizations if you choose.

```{r}
population_2020 <- population %>%
  filter(year == 2020) 



pop_2020_let <- cbind(letter, population_2020) 

p <- ggplot(pop_2020_let, aes(x = tdr, y = letter, group = letter)) +
  geom_point(size = 1) +
  geom_point(color = "blue", alpha = .2) +
  geom_text_repel(aes(label = country), size = 2)
p

```